---
title: "Distance between TURFs"
author: "Juan Carlos Villaseñr-Derbez"
output: github_document
---

# Intro

The previous Rmd  file created a figure to show the location of nine TURFs that are part of FEDECOOP. This Rmd document will now calculate the pairwise distances between all TURFs.

## Set up

```{r, include = F}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Load packages

```{r}
suppressPackageStartupMessages({
  library(here)
  library(startR)
  library(sf)
  library(SpatialPosition)
  library(tidyverse)
})
```

### Load data

```{r}
turfs <- st_read(here("raw_data", "spatial",  "feddecoop_polygons.gpkg")) %>% 
  st_centroid()
```

# Calculations

## Calculate distance between TURF centroids


We can now calculate the distance between all combinations of TURFs. This will produce a 9 X 9 matrix, with a diagonal with value so 0.

```{r}
# Assign rownames based on TURF name
rownames(turfs) <- turfs$Coop

# Calculate distance matrix (results are in meters)
matrix <- round(
  CreateDistMatrix(knownpts = turfs, unknownpts = turfs) / 1e3, # Divide by a thousand to convert to km
  2)                                                            # Keep only two decimals
```


```{r}
knitr::kable(matrix)
```

Now, lets convert the matrix into a table just in case this is useful.

```{r}
distance_as_table <- matrix %>%
  as_tibble() %>% 
  mutate(from = colnames(.)) %>% 
  gather(to, value, -from) %>% 
  mutate(value = value)
```

Now let's visualize and export this matrix.

```{r, fig.height = 7, fig.width = 7}
heatmap <- ggplot(data = distance_as_table,
       mapping = aes(x = from, y = to, fill = value)) +
  geom_tile(color = "black", size = 0.5) +
  coord_equal() +
  ggtheme_plot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = guide_colorbar(ticks.colour = "black",
                               frame.colour = "black")) + 
  labs(x = "", y = "") +
  scale_fill_viridis_c(name = "Distance (km)") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))

heatmap
```

We can now export both, the matrix and table, as well as the figure.

```{r}
write.csv(x = matrix,
          file = here::here("results", "distance_matrix.csv"))

write.csv(x = distance_as_table,
          file = here::here("results", "distance_table.csv"),
          row.names = F)

lazy_ggsave(plot = heatmap,
            filename = "distance_heatmap",
            width = 17,
            height = 17)
```


















