---
title: "Abalone Fishery"
author: "Juan Carlos Villaseñr-Derbez"
output: github_document
---

# Intro

This documents uses landings data from CONAPESCA (2000 - 2014). I will create a monthly time series of landings and values for each cooperative.

## Set up

```{r, include = F}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

### Load packages

```{r}
suppressPackageStartupMessages({
  library(here)
  library(startR)
  library(rnaturalearth)
  library(sf)
  library(janitor)
  library(tidyverse)
})
```

### Load data

```{r}
conapesca <- readRDS(here("raw_data", "conapesca.rds")) %>% 
  clean_names()

turfs <- st_read(here("raw_data", "spatial",  "fedecoop_polygons.gpkg"),
                 stringsAsFactors = F) %>% 
  st_transform("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84")
```

## Prepare data

Thes are some very messy data. We'll first have to filter it to keep only the nince cooperatives that are part of FEDECOOP. Then, we'll keep only data on abalone, and then filter again to keep only data on abalone where weight is reported without accounting for abalone shell. Then I will modify the coperative names so that they match the spatial dataset.

Once all the filters and modifications are ready, I'll then group data at the coop-year level and create a time series of landings. Landings data will be reported in metric tones. The resulting table contins the total annual landings at the coop level. A preview of this panel is this:

```{r}
# A list of the cooperatives I want
coops_i_want <- c("Scpp bahia tortugas s.c de r.l", "Scpp bahia tortugas s.c de r.l", "Scpp buzos y pescadores de la baja california scl", "Scpp california de san ignacio scl", "Scpp emancipacion sc de rl", "Scpp la purisima scl", "Soc coop de prod pesq rib leyes de reforma sc de rl", "Scpp progreso scl", "Scpp punta abreojos scl", "Scpp punta abreojos sc de rl", "Scpp pesc nacionales de abulon sc de rl")

# A list of the valid entries for abalone
abalone_types_i_want <- c(
  # "Abulon amarillo s.c. fco.",
  # "Abulon amarillo ent. fco.",
  # "Abulon chino s.c. fco.",
  # "Abulon chino ent. fco.",
  # "Abulon rayado ent. fco.",
  # "Abulon rojo s.c. fco.",
  # "Abulon rojo ent. fco.",
  "Abulon azul s.c. fco.",
  # "Abulon azul ent. fco.",
  # "Abulon blanco s.c. fco.",
  # "Abulon blanco ent. fco.",
  # "Abulon negro s.c. fco.",
  # "Abulon negro ent. fco.",
  "Abulon s.c. fco.",
  "Abulon ent. fco."
  )

# A clearn list of cooepratives, so that databases match between them
coop_dictionary <- tibble(
  coop = turfs$coop,
  unidad_economica = c("bahia tortugas",
                       "buzos y pescadores la baja california",
                       "california san ignacio",
                       "emancipacion",
                       "pesc nacionales abulon",
                       "la purisima",
                       "soc leyes reforma",
                       "progreso",
                       "punta abreojos"))

# Now, filter, clean, and group by
abalone_fishery <- conapesca %>% 
  filter(unidad_economica %in% coops_i_want) %>% 
  filter(str_detect(nombre_principal, "Abulon")) %>%
  filter(ano <= 2014) %>% 
  filter(nombre_comun %in% abalone_types_i_want) %>% 
  mutate(peso = peso_vivo) %>% 
  # mutate(tipo = ifelse(str_detect(nombre_comun, "s.c."), "sin_concha", "entero")) %>%
  select(unidad_economica,
         mes,
         year = ano,
         landed_weight = peso) %>% 
  mutate(unidad_economica = tolower(unidad_economica),
         unidad_economica = str_remove_all(unidad_economica, "scpp "),
         unidad_economica = str_remove_all(unidad_economica, "coop de prod pesq rib"),
         unidad_economica = str_remove_all(unidad_economica, "de"),
         unidad_economica = str_remove_all(unidad_economica, " scl"),
         unidad_economica = str_remove_all(unidad_economica, " sc"),
         unidad_economica = str_remove_all(unidad_economica, " rl"),
         unidad_economica = str_remove_all(unidad_economica, " s.c."),
         unidad_economica = str_remove_all(unidad_economica, " r.l."),
         unidad_economica = str_remove_all(unidad_economica, " r.l"),
         unidad_economica = str_trim(unidad_economica),
         unidad_economica = str_squish(unidad_economica)) %>% 
  left_join(coop_dictionary, by = "unidad_economica") %>% 
  select(coop, year, landed_weight) %>% 
  group_by(year, coop) %>% 
  summarize(landed_weight = sum(landed_weight, na.rm = T) / 1000) %>% 
  ungroup() %>% 
  arrange(year, coop)
```

```{r}
head(abalone_fishery, 24) %>% 
  knitr::kable(caption = "Landings (kg), revenue (MXP), and prices (MXP / Kg) for the abalone fishery in 9 cooperatives belonging to FEDECOOP.")
```

# Tables and Figures

## Total landings

Let's create a table that shows the total landings for each cooperative on any given year. We'll also add row total and column totals.

```{r}
abalone_fishery %>% 
  group_by(year, coop) %>% 
  summarize(landings = sum(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  spread(coop, landings) %>% 
  adorn_totals(where = c("row", "col"), na.rm = T) %>% 
  knitr::kable(caption = "Total landings (Kg) for each cooperative through time. NAs represent missing data for a given year / cooperative combination.")
```

The table above is shown in the figures below. The first figure shows the timseries of landed abalone for each cooperative. The second figure shows the aggregate abalone landings for all FEDECOOP cooperatives.

```{r}
abalone_fishery %>% 
  ggplot(mapping = aes(x = year, y = landed_weight, color = coop, fill = coop)) +
  stat_summary(geom = "line", fun = sum) +
  scale_color_brewer(name = "Cooperative", palette = "Set1") +
  ggtheme_plot() +
  ggtitle("Total annual landings for each cooperative") +
  labs(x = "Year", y = "Landings (Tones)")
```


```{r}
ts <- abalone_fishery %>% 
  group_by(year) %>% 
  summarize(landed_weight = sum(landed_weight, na.rm = T)) %>% 
  ggplot(aes(x = year, y = landed_weight)) +
  geom_hline(aes(yintercept = mean(landed_weight, na.rm = T) - sd(landed_weight, na.rm = T))) +
  geom_hline(aes(yintercept = mean(landed_weight, na.rm = T)), linetype = "dashed") +
  geom_hline(aes(yintercept = mean(landed_weight, na.rm = T) + sd(landed_weight, na.rm = T))) +
  geom_line(color = "black") +
  geom_point(color = "black", fill = "steelblue", shape = 21, size = 2) +
  ggtitle("Total landings by all FEDECOOP coperatives") +
  labs(x = "Year", y = "Landings (Tones)") +
  ggtheme_plot()

ts

lazy_ggsave(plot = ts,
            filename = "landings_timeseries",
            width = 10,
            height = 5)
```

## Mean measures

The tables and figures above show the totals for each year and across all cooperatives. Let's create one where we show the mean annual landings and revenue for each cooperative, and then one where we look at the grand mean.

```{r}
abalone_fishery %>% 
  ungroup() %>% 
  select(-year) %>% 
  group_by(coop) %>% 
  summarize(mean = mean(landed_weight, na.rm = T),
            sd = sd(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  knitr::kable(caption = "Annual Mean and SD landed weigh for each cooperative.")
```

```{r}
abalone_fishery %>% 
  ungroup() %>% 
  select(-year) %>% 
  group_by(coop) %>% 
  summarize(mean = mean(landed_weight, na.rm = T),
            sd = sd(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = coop)) + 
  geom_errorbar(aes(ymin = mean, ymax = mean + sd, width = 0.2)) +
  geom_col(aes(y = mean), fill = "steelblue", color = "black") +
  ggtheme_plot() +
  coord_flip()
```


```{r, fig.width = 5, fig.height = 5}
baja <- ne_states(country = "Mexico", returnclass = "sf") %>% 
  filter(name %in% c("Baja California", "Baja California Sur")) %>% 
  st_crop(st_buffer(st_as_sfc(st_bbox(turfs)), 1)) %>% 
  st_union()

a <- abalone_fishery %>% 
  group_by(coop) %>% 
  summarize(landed_weight = mean(landed_weight, na.rm = T))

turfs %>% 
  left_join(a) %>% 
  ggplot() +
  geom_sf(data = baja, color = "black") +
  geom_sf(aes(fill = landed_weight)) +
  ggtheme_map() +
  scale_fill_viridis_c(name = "Mean landings\n2000-2014\n(metric tones)") +
  guides(fill = guide_colorbar(frame.colour = "black",
                               ticks.colour = "black"))
```

```{r, fig.width = 5, fig.height = 5}
turfs %>% 
  mutate(area = st_area(.),
         area = as.numeric(area) / 1e6) %>% 
  left_join(a) %>% 
  mutate(landed_weight_norm = landed_weight / area) %>% 
  ggplot() +
  geom_sf(data = baja, color = "black") +
  geom_sf(aes(fill = landed_weight_norm)) +
  ggtheme_map() +
  scale_fill_viridis_c(name = "Mean landings\n2000-2014\n(tones / Km2)") +
  guides(fill = guide_colorbar(frame.colour = "black",
                               ticks.colour = "black"))
```

```{r}

grand_mean <- abalone_fishery %>% 
  group_by(year) %>% 
  summarize(weight = sum(landed_weight, na.rm = T)) %>% 
  pull(weight)

gm <- mean(grand_mean)
gsd <- sd(grand_mean)

```

Based on the 2000 - 2014 data, a cooperative's mean catch is `r mean(abalone_fishery$landed_weight)` $\pm$ `r sd(abalone_fishery$landed_weight)` tones (mean $\pm$ SD). Total annual extraction to be `r gm` $\pm$ `r gsd` (mean $\pm$ SD). In FEDECOOPs website (https://www.fedecoop.com.mx/), they report that current total abalone extraction averages 290 metric tonnes for all cooperatives together. This value is closer to the more recent value in these data.


However, this analyses des not accout for the temporal trend in which cooperatives have been forced to reduce fishing effort due to the diwndling status of stocks. I





































