---
title: "Abalone Fishery"
author: "Juan Carlos Villaseñr-Derbez"
output: github_document
---

# Intro

This documents uses landings data from CONAPESCA (2000 - 2014). I will create a monthly time series of landings and values for each cooperative.

## Set up

```{r, include = F}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

### Load packages

```{r}
suppressPackageStartupMessages({
  library(here)
  library(startR)
  library(rnaturalearth)
  library(sf)
  library(janitor)
  library(tidyverse)
})
```

### Load data

```{r}
conapesca <- readRDS(here("raw_data", "conapesca.rds")) %>% 
  clean_names()

turfs <- st_read(here("raw_data", "spatial",  "fedecoop_polygons.gpkg"),
                 stringsAsFactors = F) %>% 
  st_transform("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84")
```

## Prepare data

This is some very messy data. We'll first have to filter to include only the nince cooperatives that are part of FEDECOOP. Then, we'll keep only data on abalone, and then filter again to keep only data on abalone where weight is reported without accounting for abalone shell. Then I will modify the coperative names so that they match the spatial dataset.

Once all teh filters and modifications are ready, I'll then group data at the coop-year level and create a timeseries of landings and value associated with those landings.

```{r}
# A list of the cooperatives I want
coops_i_want <- c("Scpp bahia tortugas s.c de r.l", "Scpp bahia tortugas s.c de r.l", "Scpp buzos y pescadores de la baja california scl", "Scpp california de san ignacio scl", "Scpp emancipacion sc de rl", "Scpp la purisima scl", "Soc coop de prod pesq rib leyes de reforma sc de rl", "Scpp progreso scl", "Scpp punta abreojos scl", "Scpp punta abreojos sc de rl", "Scpp pesc nacionales de abulon sc de rl")

# A list of the valid entries for abalone
abalone_types_i_want <- c(
  "Abulon amarillo s.c. fco.",
  # "Abulon amarillo ent. fco.",
  "Abulon chino s.c. fco.",
  # "Abulon chino ent. fco.",
  # "Abulon rayado ent. fco.",
  "Abulon rojo s.c. fco.",
  # "Abulon rojo ent. fco.",
  "Abulon azul s.c. fco.",
  # "Abulon azul ent. fco.",
  "Abulon blanco s.c. fco.",
  # "Abulon blanco ent. fco.",
  "Abulon negro s.c. fco.",
  # "Abulon negro ent. fco.",
  "Abulon s.c. fco.",
  "Abulon ent. fco."
  )

# A clearn list of cooepratives, so that databases match between them
coop_dictionary <- tibble(
  coop = turfs$coop,
  unidad_economica = c("bahia tortugas",
                       "buzos y pescadores la baja california",
                       "california san ignacio",
                       "emancipacion",
                       "pesc nacionales abulon",
                       "la purisima",
                       "soc leyes reforma",
                       "progreso",
                       "punta abreojos"))

# Now, filter, clean, and group by
abalone_fishery <- conapesca %>% 
  filter(unidad_economica %in% coops_i_want) %>% 
  filter(str_detect(nombre_principal, "Abulon")) %>%
  filter(nombre_comun %in% abalone_types_i_want) %>% 
  mutate(peso = peso_vivo) %>% 
  # mutate(tipo = ifelse(str_detect(nombre_comun, "s.c."), "sin_concha", "entero")) %>%
  select(unidad_economica,
         mes,
         year = ano,
         landed_weight = peso) %>% 
  mutate(unidad_economica = tolower(unidad_economica),
         unidad_economica = str_remove_all(unidad_economica, "scpp "),
         unidad_economica = str_remove_all(unidad_economica, "coop de prod pesq rib"),
         unidad_economica = str_remove_all(unidad_economica, "de"),
         unidad_economica = str_remove_all(unidad_economica, " scl"),
         unidad_economica = str_remove_all(unidad_economica, " sc"),
         unidad_economica = str_remove_all(unidad_economica, " rl"),
         unidad_economica = str_remove_all(unidad_economica, " s.c."),
         unidad_economica = str_remove_all(unidad_economica, " r.l."),
         unidad_economica = str_remove_all(unidad_economica, " r.l"),
         unidad_economica = str_trim(unidad_economica),
         unidad_economica = str_squish(unidad_economica)) %>% 
  left_join(coop_dictionary, by = "unidad_economica") %>% 
  select(coop, year, landed_weight) %>% 
  group_by(year, coop) %>% 
  summarize(landed_weight = sum(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  arrange(year, coop)
```

The resulting table contins the total monthly landings and revenue (as well as price) at the coop-year-month level. A preview of this panel is this:

```{r}
head(abalone_fishery, 24) %>% 
  knitr::kable(caption = "Landings (kg), revenue (MXP), and prices (MXP / Kg) for the abalone fishery in 9 cooperatives belonging to FEDECOOP.")
```

# Tables and Figures

## Total landings and revenues

Let's create a table that shows the total landings for each cooperative on any given year.

```{r}
abalone_fishery %>% 
  group_by(year, coop) %>% 
  summarize(landings = sum(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  spread(coop, landings) %>% 
  adorn_totals(where = c("row", "col"), na.rm = T) %>% 
  knitr::kable(caption = "Total landings (Kg) for each cooperative through time. NAs represent missing data for a given year / cooperative combination.")
```


```{r}
abalone_fishery %>% 
  ggplot(mapping = aes(x = year, y = landed_weight, color = coop, fill = coop)) +
  stat_summary(geom = "line", fun = sum) +
  scale_color_brewer(name = "Cooperative", palette = "Set1") +
  ggtheme_plot() +
  ggtitle("Total annual landings and revenue for each cooperative") +
  labs(x = "Year", y = "value (in Kg or MXP)")
```


```{r}
abalone_fishery %>% 
  group_by(year) %>% 
  summarize(landed_weight = sum(landed_weight, na.rm = T)) %>% 
  ggplot(aes(x = year, y = landed_weight)) +
  geom_line(color = "black") +
  geom_point(color = "black", fill = "steelblue", shape = 21, size = 2) +
  ggtitle("Total landings by all FEDECOOP coperatives") +
  labs(x = "Year", y = "value (in Kg or MXP)") +
  ggtheme_plot()
```

## Mean measures

The tables and figures above show the totals for each year and across all cooperatives. Let's create one where we show the mean annual landings and revenue for each cooperative, and then one where we look at the grand mean.

```{r}
abalone_fishery %>% 
  ungroup() %>% 
  select(-year) %>% 
  group_by(coop) %>% 
  summarize(mean = mean(landed_weight, na.rm = T),
            sd = sd(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  knitr::kable(caption = "Annual Mean and SDlanded weigh for each cooperative.")
```

```{r}
abalone_fishery %>% 
  ungroup() %>% 
  select(-year) %>% 
  group_by(coop) %>% 
  summarize(mean = mean(landed_weight, na.rm = T),
            sd = sd(landed_weight, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = coop)) + 
  geom_errorbar(aes(ymin = mean, ymax = mean + sd, width = 0.2)) +
  geom_col(aes(y = mean), fill = "steelblue", color = "black") +
  ggtheme_plot() +
  coord_flip()
```


```{r, fig.width = 5, fig.height = 5}
baja <- ne_states(country = "Mexico", returnclass = "sf") %>% 
  filter(name %in% c("Baja California", "Baja California Sur")) %>% 
  st_crop(st_buffer(st_as_sfc(st_bbox(turfs)), 1)) %>% 
  st_union()

a <- abalone_fishery %>% 
  group_by(coop) %>% 
  summarize(landed_weight = mean(landed_weight, na.rm = T))

turfs %>% 
  left_join(a) %>% 
  ggplot() +
  geom_sf(data = baja) +
  geom_sf(aes(fill = landed_weight / 1e3)) +
  ggtheme_map() +
  scale_fill_viridis_c(name = "Mean landings\n2000-2014\n(metric tones)") +
  guides(fill = guide_colorbar(frame.colour = "black",
                               ticks.colour = "black"))
```

```{r, fig.width = 5, fig.height = 5}
turfs %>% 
  mutate(area = st_area(.),
         area = as.numeric(area) / 1e6) %>% 
  left_join(a) %>% 
  mutate(landed_weight_norm = (landed_weight / 1000) / (area / 1000)) %>% 
  ggplot() +
  geom_sf(data = baja) +
  geom_sf(aes(fill = landed_weight_norm)) +
  ggtheme_map() +
  scale_fill_viridis_c(name = "Mean landings\n2000-2014\n(metric tones / Km2)") +
  guides(fill = guide_colorbar(frame.colour = "black",
                               ticks.colour = "black"))
```


However, this analyses des not accout for the temporal trend in which cooperatives have been forced to reduce fishing effort due to the diwndling status of stocks. n FEDECOOPs website (https://www.fedecoop.com.mx/), they report that total extraction averages 290 metric tonnes for all cooperatives together. This value is closer to the more recent values.





































