---
title: "Abalone model"
author: "Juan Carlos Villaseñr-Derbez"
output: github_document
---

# Intro

This documents constructs a spatially-explicit and size-structured model for Abalones in Marine Reserve contexts. This is a reconstru tion of the 2015 paper by Marissa Rosetto: 

> Rossetto, M., Micheli, F., Saenz-Arroyo, A., Montes, J. A. E., & De Leo, G. A. (2015). No-take marine reserves can enhance population persistence and support the fishery of abalone. Canadian Journal of Fisheries and Aquatic Sciences, 72(10), 1503-1517.

## Set up

```{r, include = F}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

### Load packages

```{r}
suppressPackageStartupMessages({
  library(here)
  library(startR)
  library(tidyverse)
})
```

# Model

The model has 150 patches, each of which represent a 100 X 500 m parcel (*i.e.* 5 hectares). The population contains 9 size classes, in 25 mm increments. Density-dependence occurs during settlement of larvae.

## Define model parameters and components

All these parameters are presented roughly in the same order as Table 1 in Rosstto et al 2015.


### Index and spatial variables

```{r}
n_steps <- 150                                             # Number of iterations into the future
n_classes <- 9                                            # Number of size classes
n_patches <- 150                                          # Number of patches in the system
A <- 5                                                    # Patch area (5 hectares)

N <- array(
  dim = c(n_classes,                                      # Each row in a matrix is a size
          n_patches,                                      # Each column in a matrix is a patch
          n_steps                                         # Each layer in the array is a time step
  ))

N[, , 1] <- c(1000, rep(10, 8))                                           # Set initial population sizes fo all classes and patches
r <- rep(c(0, 1, 1, 0, 1, 1, 1, 0, 1, 1), 15)                                  # Reserve system
```

### Lenght and mass

```{r}
size_class <- c("5–30",                                 # Size classes as text
                "30–55",
                "55–80",
                "80–105",
                "105–130",
                "130–155",
                "155–180",
                "180–205",
                "205–230")
l_i <- 17.5 + (25 * 0:(n_classes - 1))                  # Mean abalone length for each length class
a <- 2.24e-5                                            # "a" parameter in allometric growth
b <- 3.36                                               # "b" paramteter in allometric growth
w_i <- a * (l_i ^ b)                                    # Weight-at-length for each class
```

### Harvest

```{r}
# Fixed parameters
mls <- 155                                               # Minimum Landings Size (155 mm)
h <- 0.7                                                 # Harvest rate (0, 1). This should be the control variable
h_i <- (l_i >= mls) * h                                  # Vector of size-specific harvesting
h_iz <- h_i %*% t(r)                                     # Matrix that indicates where harvesting occurs for esch size class

# A function to talculate harvest
catch <- function(w_i, h, N_izt){
  sum((w_i * h * N_izt * 5) * 10e6)
}
```

### Natural Mortality in adults

In abalone, natural mortality rates of settled individuals have been shown to be size-dependent, with survival probabilities increasing with size (Rossetto et al. 2012). Accordingly, size-specific mortality rates $\mu_i$ (year−1) were calculated from mean body mass in each size class wi (g) using the empirical allometric relationship between instantaneous mortality rates and body mass reported for abalones in natural environments (Rossetto et al. 2012):

$$
ln(\mu_i) = \theta + \alpha ln(l_i)
$$
Annual survival si (i.e., the fraction of individuals in class i that survive to the following year) was then computed as:

$$
s_i = e^{-\mu_i}
$$

```{r}
theta_m <-  0.631
theta_sd <- 0.102
alpha_m <- -0.317
alpha_sd <- 0.027

survival <- function(theta_m, theta_sd, alpha_m, alpha_sd) {
  
  theta <- rnorm(1, mean = theta_m, sd = theta_sd)
  alpha <- rnorm(1, mean = alpha_m, sd = alpha_sd)
  
  mu_i <- exp(theta + (alpha * log(w_i)))
  exp(theta) * w_i ^ alpha
  s_i <- exp(-mu_i)
  
  return(s_i)
}
```

### Somatic growth

From Table A1 - "Growth transition g_{i,j}, from initial class j to final class i

```{r}
# Create matrix
G <- matrix(
  data = c(0.166, 0, 0, 0, 0, 0, 0, 0, 0,
           0.590, 0.02, 0, 0, 0, 0, 0, 0, 0,
           0.242, 0.533, 0.014, 0, 0, 0, 0, 0, 0,
           0.003, 0.429, 0.553, 0.028, 0, 0, 0, 0, 0,
           0, 0.018, 0.424, 0.77, 0.102, 0, 0, 0, 0,
           0, 0, 0.008, 0.202, 0.878, 0.317, 0, 0, 0,
           0, 0, 0, 0, 0.02, 0.683, 0.549, 0, 0,
           0, 0, 0, 0, 0, 0, 0.451, 0.694, 0,
           0, 0, 0, 0, 0, 0, 0, 0.306, 1),
  nrow = 9,
  byrow = TRUE )

# Assign names
colnames(G) <- size_class
rownames(G) <- size_class

G
```

### Fecundity

```{r}
# Fixed parameters
e_mean <- 3772                                          # Mean umber of eggs produced per gram of female individual
e_sd <- 330                                             # SD of the above
nu <- 135.99                                            # inflection point from Rosetto et al., 2013
mu <- 20.20                                             # inv. prop. to the slope at 1=mu From Rosetto et al., 2013
psi_i <- 1 / (1 + exp(-((l_i - nu)/(mu))))              # Fraction of sexually mature individuals of class i      

# The vector of number of eggs produced in each patch is produced by this function, which depends on:
# - eggs produced per gram of female individual
# - mean weight of individuals in class i
# - fraction of sexualy matyre individuals of class i
# - number of organisms in each patch and size class
eggs <- function(e, w_i, psi_i, N_izt) {
  E <- 0.5 * e * colSums(w_i * psi_i * N_izt)
  return(E)
}
```

### Larval and settler survival

```{r}
# Fixed parameters
K <- 10e7                                               # Carrying capacity of settlers (ind / hectare)
sigma_e_m <- 3.09e-3                                    # Mean survival from eggs to settlers
sigma_e_sd <- 0.33e-3                                   # SD of survival
sigma_o <- 0.01                                         # Survival of settlers at low densities

# Functions that create parameters
# Survival of settlers during the first year of life depends on:
# - the survival of settlers at low densities
# - number of settlers
# - Carrying capacity
sigma_s <- function(sigma_o, S, K) {
  sigma_o * exp((-S/(K)))
}
```

### Larval dispersal

```{r, fig.height = 2, fig.width = 4}
d <- 300                                                # Larval dispersal distance (m)
sigma_d <- 116.4674                                     # Set so that 99% of larvae are retained at dist =/- d
pnorm(d, 0, sd = sigma_d) - pnorm(-d, 0, sd = sigma_d)  # Proportion of larvae retained?

ks <- (1:150) - 75                                      # Vector of patch indices
xs_u <- 100 * (ks + 0.5)                                # Vector fo patch distances (upper integral limit)
xs_l <- 100 * (ks - 0.5)                                # Vector of patch distances (lower integral limit)
rho <- pnorm(xs_u, 0, sd = sigma_d) - pnorm(xs_l, 0, sd = sigma_d)# Proportion of larvae ending up in each patch
plot(ks, rho, type = "l")
```

For simplicity, I'll create a connectivity matrix.

```{r}
rho_mod <- c(rho[75:n_patches], rho[1:74])                # Modify indices
rho_mat <- matrix(nrow = n_patches,                       # Create the 9 X 9 dispersal matrix
                  ncol = n_patches)
rho_mat[1,] <- rho_mod                                    # First row is simply the rearranged one
# Iterate across i (patch) to build the matrix
for(i in 2:n_patches) {
  rho_i <- c(rho_mod[(n_patches - i + 2) : n_patches], rho_mod[1:(n_patches - i + 1)])
  rho_mat[i, ] <- rho_i
}

image(rho_mat)
```

Then, settlers are just a function of the number of eggs produced, and where they end up in

```{r}
settlers <- function(sigma, E, rho_mat) {
  sigma * colSums(E *  rho_mat)
}
```


## Building the model


### Transition Matrix $M$

The transition matrix needs three pieces of information. Growth, survival, and harvest.

```{r, eval = F}
M <- s_i * (1 - h_i) * G                                 # The M matrix is defined in equation 2 of the paper
M
```


```{r}

n_t_i <- matrix(numeric(length = n_steps * 1000), nrow = n_steps)

for (j in 1:1000){
  for(i in 2:n_steps) {
    # Eggs
    ## Egg production
    e <- rnorm(1, mean = e_mean, sd = e_sd)
    E <- eggs(e = e,
              w_i = w_i,
              psi_i = psi_i,
              N_izt = N[, , i-1])
    ## Egg survival into settlers
    sigma_e <- rnorm(1, mean = sigma_e_m, sd = sigma_e_sd)
    S <- settlers(sigma = sigma_e,
                  E = E,
                  rho_mat = rho_mat)
    
    ## Adult survival
    s_i <- survival(theta_m, theta_sd, alpha_m, alpha_sd)
    M <- s_i * G 
    
    ss <- sigma_s(sigma_o = sigma_o,
                  S = S,
                  K = K)
    R <- ss * S
    
    N_after_harvest <- (1 - h_iz) * N[, , i-1]
    N_now <- M %*% N_after_harvest
    N_now[1, ] <- N_now[1, ] + R
    
    N[, , i] <- N_now
  }
  
  n_t <- apply(N[-1,,], 3, sum, na.rm = T)
  n_t_i[ , j] <- n_t
}
```

```{r}
n_t <- apply(n_t_i, 1, mean)

matplot(n_t_i, type = "l", lwd = 0.1, pch = ".")
lines(n_t, type = "l", col = "black")
abline(h = mean(n_t[-c(1:50)]), col = "blue")

mean(n_t[-c(1:50)]) / 1e6
```

```{r}
nti_df <- n_t_i %>% 
  magrittr::set_colnames(value = paste("try", 1:1000, sep = "_")) %>% 
  as_tibble() %>% 
  mutate(t = 1:n_steps) %>% 
  gather(try, n, -t) %>% 
  mutate(n = n / 1e6)

ggplot(data = nti_df, aes(x = t, y = n)) +
  stat_summary(geom = "ribbon", fun.data = mean_sdl, fill = "transparent", color = "black", linetype = "dashed") +
  geom_line(aes(group = try), alpha = 0.5, size = 0.01, color = "steelblue") +
  stat_summary(geom = "line", fun = "mean") +
  labs(title = "1000 model runs", x = "Time", y = "N_t (millions)") +
  ggtheme_plot()
```



```{r}
plot(n_t, c(NA, diff(n_t)), main = "State-State plot", xlab = "N_t", ylab = "N_t+1")
```


```{r}
get_r <- function(x) {
  mean(log(x / lag(x)), na.rm = T)
}

rs <- apply(n_t_i, 2, get_r)

plot(density(rs), "Density distribution of r")
abline(v = mean(rs))
text(x = mean(rs), y = 100, paste("Mean is ", round(mean(rs), 4)))


r_dens <- group_by(nti_df, try) %>% 
  summarize(r = get_r(n)) %>% 
  ggplot(aes(x = r)) +
  geom_density(fill = "steelblue") +
  ggtheme_plot() +
  geom_vline(aes(xintercept = mean(r)), linetype = "dashed") +
  geom_text(aes(x = mean(r) + sd(r), y = 100, label = paste("Mean is ", round(mean(rs), 4)))) +
  labs(title = "Density distribution of r", x = "r", y = "Density")

lazy_ggsave(r_dens, filename = "r_dens", width = 10, height = 6)
```




